---
title: 'Exercise Sheet #3'
subtitle: "Fortgeschrittene Statistische Software für NF"
author: "Minh Son Tran (12424799)"
date: "`2025-06-05"
#output: 
#  bookdown::pdf_document2: default
output: distill::distill_article
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.align  = "center"
)

```

```{r, echo = FALSE, message=FALSE, warning=FALSE}
#install.packages("tidyverse")
#install.packages("palmerpenguins")
#install.packages("dplyr")
#install.packages("knitr")
#install.packages("easystats")
#install.packages("stargazer")
#install.packages("margins")
#install.packages("reticulate")
#install.packages("effects")
#install.packages("carData")
#install.packages("sf")
library(readr)
library(tidyverse)
library(palmerpenguins)
library(dplyr)
library(knitr)
library(ggplot2)
library(report)
library(parameters)
library(see)
library(reticulate)
library(margins)
library(lattice)
library(effects)
library(sf)
```

## Exercise 1

### d)

Git's Strengths and Weaknesses:

2 strengths:

-   It allow user to create branches, makes it easy to experiment
    without affecting the main codebase.

-   It is open source and actively maintained.

2 weaknesses:

-   User requires time to learn how to use Git properly because of
    confusing commands.

-   Git is not very good for project with binary or very large files

## Exercise 2

### a)

GitHub repo: <https://github.com/MinhSonTran97/exeRcise-sheet-3.git>

## Exercise 3

### a)

```{r echo = TRUE, message=FALSE, warning=FALSE}
pixar_films <- read_csv("data/pixar_films.csv") %>%
  filter(!is.na(film))
```

The possible film_rating values are `r unique(pixar_films$film_rating)`:

-   "G": General audiences

-   "PG": Parental guidance is advised

-   "N/A": Film Rating is not available / no rating provided. This is
    included since we only removed entries with missing titles.

```{r echo = TRUE, message=FALSE, warning=FALSE}
pixar_films <- pixar_films %>%
  mutate(film_rating = factor(film_rating, levels = c("G", "PG"), ordered = TRUE))
```

Converting film_rating into rating_factor is appropriate because:

-   film_rating represents a set of discrete categories, not numeric
    values. These categories describe the type of audience the film is
    suitable for, not a measurable quantity.

-   Treating film_rating as a factor helps R recognize it as a
    categorical variable, which allow us to better summarize, plot, and
    handle the data in modeling.

### b)

```{r echo = TRUE, message=FALSE, warning=FALSE}
pixar_films <- pixar_films %>%
  mutate(series = case_when(
    str_detect(film,"Toy Story") ~ "Toy Story",
    str_detect(film,"Cars") ~ "Cars",
    str_detect(film,"Incredibles") ~ "Incredibles",
    str_detect(film,"Finding") ~ "Finding",
    str_detect(film,"Monsters") ~ "Monsters",
    TRUE ~ NA_character_
  )) %>%
  group_by(series)

film_series <- pixar_films %>%
  filter(str_detect(film,"Toy Story|Cars|Incredibles|Finding|Monsters")) %>%
  summarise(films = paste(film, collapse = " - "), number_of_entries = n()) %>%
  ungroup()
film_series
```

### c)

```{r echo = TRUE, message=FALSE, warning=FALSE}
public_response <- read_csv("data/public_response.csv")

unique(public_response$cinema_score)  
```

We can see that the only rating scores are
`r unique(public_response$cinema_score)`. We would assume that NA means
the score is missing or not included, meaning the scoring system will
look like this : A- \< A \< A+

```{r echo = TRUE, message=FALSE, warning=FALSE}
public_response <- public_response %>%
  mutate(cinema_score = factor(cinema_score, levels = c("A-", "A", "A+"), ordered = TRUE))

pixar_sum <- left_join(pixar_films, public_response, by = "film")
```

### d)

```{r echo = TRUE, message=FALSE, warning=FALSE}
pixar_series <- pixar_sum %>%
  filter(!is.na(series)) %>%
  filter(!is.na(metacritic))
```

```{r pixar-metacritic, fig.cap = "Performance comparison of sequels versus their first installments in film series", echo = TRUE, message=FALSE, warning=FALSE}
ggplot(
  data = pixar_series,
  aes(
    x = series,
    y = metacritic,
    fill = fct_reorder(film, release_date)
  )) +
  geom_col(
    width = 0.75,
    position = position_dodge(width = 0.75)) +
  labs(
    x = NULL,
    y = "Metacritic",
    fill = "Film (chronological)",
    title = "Sequel vs. Original: Metacritic Comparison"
    ) +
  theme_minimal()
```

We can see in Figure \@ref(fig:pixar-metacritic) that the first
installments in each series all have the highest Metacritic score. After
that there, is a clear downward trend in most series, meaning that the
sequels usually couldn't perform as well as the prequels.

## Exercise 4

### c)

-   Dataset Name: Cross-country literacy rates - World Bank, CIA World
    Factbook, and other sources
-   Link:
    <https://github.com/owid/owid-datasets/tree/master/datasets/Cross-country%20literacy%20rates%20-%20World%20Bank%2C%20CIA%20World%20Factbook%2C%20and%20other%20sources>

### d)

Research question: How has the global literacy rate changed between 1970
and 2015?

The period from 1970 to 2015 includes major global education
initiatives, such as UNESCO’s Education for All, making it an
interesting and relevant timeframe to assess how these efforts have
impacted the average literacy rate worldwide.

### e)

```{r echo = TRUE, message=FALSE, warning=FALSE}
world_literacy <- read_csv("data/Cross-country literacy rates - World Bank, CIA World Factbook, and other sources.csv")

literacy_trend <- world_literacy %>%
  filter(Year >= 1970) %>%
  group_by(Year) %>%
  summarise(avg_world_literacy = mean(`Literacy rates (World Bank, CIA World Factbook, and other sources)`), na.rm = TRUE)

literacy_change <- literacy_trend %>%
  filter(Year %in% c(1970, 2015)) %>%
  mutate(change = round(avg_world_literacy - lag(avg_world_literacy), 1)) %>%
  ungroup()
```

The global mean literacy rate rose from
`r round(literacy_change$avg_world_literacy[1],2)`% in
`r round(literacy_change$Year[1])` to
`r round(literacy_change$avg_world_literacy[2],2)`% in
`r round(literacy_change$Year[2])`, an increase of about
`r round(literacy_change$change[2],2)`% over 45 years. While the
increase may seem modest, it represents millions of additional literate
people globally, which is a significant improvement, especially in
regions starting with lower literacy.

### f)

```{r literacy-trend,fig.cap = "Global average and country-specific literacy rate trend, from 1970 to 2015.<br><span style='color:blue;font-weight:700;'>───</span><span style='font-size:10px;'>  Global average literacy rate over time  </span><span style='color:red;font-weight:700;'>───</span><span style='font-size:10px;'>  Global average literacy rate differences between 1970 and 2015</span>", echo = TRUE, message=FALSE, warning=FALSE}

ggplot() +
  geom_line(
    data = world_literacy %>%
      filter(Year >= 1970),
    aes(
      x = Year,
      y = `Literacy rates (World Bank, CIA World Factbook, and other sources)`,
      group = Entity,
      color = Entity),
    alpha = 0.5,
    size = 0.25,
    show.legend = FALSE
    ) +
  geom_line(
    data = literacy_trend,
    aes(
      x = Year,
      y = avg_world_literacy
    ),
    color = "blue",
    size = 1
  ) +
  geom_line(
    data = literacy_change,
    aes(
      x = Year,
      y = avg_world_literacy
    ),
    color = "red",
    size = 1.5
  ) +
  scale_y_continuous(labels = scales::percent_format(scale = 1)) +
  scale_x_continuous(
    breaks = seq(1970, 2015, by = 15),
    labels = scales::label_number()) +
  labs(
    x = "Year",
    y = "Literacy rate (%)") +
  theme_minimal()
```
